\section{Grounding the Evaluation: The Gain Is Invisible to a VLM}
\label{sec:grounded_eval}

The visual modalities expose the symmetric half of the framework (P3): even when grounded feedback \emph{does} fix the artifact, an ungrounded metric cannot see it.

\paragraph{The default VLM judge is structurally blind.} Our starting evaluator is a binary per-requirement rubric scored by a VLM over a screenshot. On $15$ dense academic-paper architecture figures, even with native-resolution quadrant tiling, this judge rates \textbf{$14$ of $15$} single-shot renders ``perfect.'' Yet by direct inspection they are broken---superimposed section titles, labels spilling out of boxes. The reason is mechanical: screenshots are captured at $1080$\,px, so any content overflowing the canvas is cropped \emph{off-frame before the image reaches the judge}, and small in-frame overlaps fall below its acuity. A deterministic DOM-geometry check finds only \textbf{$3$ of $15$} are actually clean (mean $1.6$ defects). The VLM is blind to precisely the defects that matter (\Cref{fig:blind}).

\begin{figure}[t]
\centering
\begin{tikzpicture}
\begin{axis}[width=0.66\linewidth,height=3.8cm,
  ybar,bar width=20pt,ymin=0,ymax=16,
  ylabel={single-shot figures (of 15)},
  symbolic x coords={VLM judge,DOM geometry},xtick=data,
  enlarge x limits=0.5,nodes near coords,nodes near coords style={font=\footnotesize},
  ymajorgrids,ytick={0,5,10,15}]
  \addplot[draw=cSS!60!black,fill=cSS!35] coordinates {(VLM judge,14)};
  \addplot[draw=cRV!60!black,fill=cRV!35] coordinates {(DOM geometry,3)};
  \legend{}
\end{axis}
\end{tikzpicture}
\caption{\textbf{The VLM judge is structurally blind.} Number of the same $15$ single-shot academic figures rated \emph{clean/perfect}: the VLM-on-a-screenshot judge passes $14$; the deterministic DOM-geometry instrument passes only $3$. The other $11$ are broken in ways cropped off-frame or below VLM acuity.}
\label{fig:blind}
\end{figure}

\paragraph{A deterministic instrument reveals a large, significant effect across four modalities.} We score four visual modalities with the same DOM-geometry+alignment instrument under one identical configuration (Sonnet~4, temperature~$0$, design-principle prompt, three seeds, propose$\to$measure$\to$feed-exact-defects-back$\to$revise, $\le3$ iterations). \Cref{tab:geometric} and \Cref{fig:geomreduction} report the result: grounded geometric feedback produces a large, significant reduction in real layout defects on \emph{all four}, with bootstrap 95\% CIs that exclude zero and improvements outnumbering regressions $\ge5{:}1$. The magnitudes track single-shot headroom---dense responsive web pages are defect-rich out of the box ($16$ defects/page across desktop and mobile), while a frontier model under a design-principle prompt already lays out many figures and slides cleanly.

\begin{table}[t]
\centering\small
\caption{Deterministic geometric-feedback results, one identical configuration across four visual modalities (Sonnet~4, $T{=}0$, $3$ seeds, alignment-inclusive reward). $\Delta$ is the mean defect reduction; CI is a paired bootstrap 95\% interval on $\Delta\%$ ($10$k resamples); $p$ is a two-sided paired sign test. Web is summed over $1920/375$ widths; animations over sampled frames.}
\label{tab:geometric}
\begin{tabular}{lcccccc}
\toprule
Modality & $n$ & SS & Rev. & $\Delta$ & 95\% CI & impr./regr.\ ($p$) \\
\midrule
Academic figures (20) & 60 & $0.57$ & $0.15$ & $-74\%$ & $[52,89]\%$ & $17/2$\;($7{\times}10^{-4}$) \\
Dense slides (12)     & 36 & $1.25$ & $0.33$ & $-73\%$ & $[45,93]\%$ & $13/1$\;($1.8{\times}10^{-3}$) \\
Web pages (20)        & 60 & $16.1$ & $8.5$  & $-47\%$ & $[30,62]\%$ & $33/2$\;($4{\times}10^{-8}$) \\
Animations (20)       & 60 & $16.9$ & $10.2$ & $-40\%$ & $[10,64]\%$ & $34/7$\;($3{\times}10^{-5}$) \\
\bottomrule
\end{tabular}
\end{table}

\begin{figure}[t]
\centering
\begin{tikzpicture}
\begin{axis}[width=0.86\linewidth,height=4.4cm,
  xbar,xmin=0,xmax=100,bar width=10pt,enlarge y limits=0.22,
  xlabel={geometric-defect reduction under grounded feedback (\%), bootstrap 95\% CI},
  symbolic y coords={Animations,Web pages,Dense slides,Academic figures},
  ytick=data,xtick={0,20,40,60,80,100},xmajorgrids,
  nodes near coords,nodes near coords style={font=\footnotesize,/pgf/number format/.cd,precision=0},
  nodes near coords align={anchor=west}]
  \addplot+[draw=cRV!60!black,fill=cRV!35,
    error bars/.cd,x dir=both,x explicit]
    coordinates {(74,Academic figures)+=(15,0)-=(22,0)
                 (73,Dense slides)+=(20,0)-=(28,0)
                 (47,Web pages)+=(15,0)-=(17,0)
                 (40,Animations)+=(24,0)-=(30,0)};
\end{axis}
\end{tikzpicture}
\caption{\textbf{Grounded geometric feedback removes real layout defects on all four visual modalities} (mean reduction; whiskers are paired-bootstrap 95\% CIs, all excluding zero). The standard VLM-on-a-screenshot metric measures none of this.}
\label{fig:geomreduction}
\end{figure}

\paragraph{An ungrounded \emph{reviewer} makes layout worse (ablation, two modalities).} Grounding matters on the feedback side too, controllably. Holding tasks and proposer fixed and varying \emph{only} the reviewer, a VLM-feedback reviewer \emph{increases} mean geometric defects on both modalities tested---slides $1.89\to2.44$ and figures $0.52\to0.62$ (regressions outnumber improvements $10$:$5$ on figures)---because, blind to the true geometry, its edits break alignment as often as they fix it. The deterministic geometric-feedback reviewer instead reduces defects on the same suites ($1.25\to0.33$ on slides, $0.57\to0.15$ on figures). The binary VLM rubric on these slides simultaneously saturates ($0.97$ mean, $29/36$ rated perfect), confirming the judge cannot even see the difference. Both halves of \Cref{fig:concept} are necessary (\Cref{fig:ablation}).

\begin{figure}[t]
\centering
\begin{tikzpicture}
\begin{axis}[name=axslide,width=0.42\linewidth,height=3.9cm,ybar,bar width=10pt,
  ylabel={mean defects},title={\footnotesize slides},ymin=0,ymax=2.9,ytick={0,1,2},
  symbolic x coords={VLM-fb,geom-fb},xtick=data,enlarge x limits=0.6,
  nodes near coords,nodes near coords style={font=\scriptsize,/pgf/number format/.cd,fixed,precision=2},ymajorgrids]
  \addplot[draw=cGrey,fill=black!10] coordinates {(VLM-fb,1.89) (geom-fb,1.25)};
  \addplot[draw=cAccent!60!black,fill=cAccent!25] coordinates {(VLM-fb,2.44) (geom-fb,0.33)};
\end{axis}
\begin{axis}[at={(axslide.right of east)},anchor=left of west,xshift=1.1cm,
  width=0.42\linewidth,height=3.9cm,ybar,bar width=10pt,
  title={\footnotesize academic figures},ymin=0,ymax=0.8,ytick={0,0.25,0.5},
  symbolic x coords={VLM-fb,geom-fb},xtick=data,enlarge x limits=0.6,
  nodes near coords,nodes near coords style={font=\scriptsize,/pgf/number format/.cd,fixed,precision=2},ymajorgrids,
  legend style={font=\footnotesize,at={(0.5,1.12)},anchor=south,legend columns=2,draw=cGrey!50}]
  \addplot[draw=cGrey,fill=black!10] coordinates {(VLM-fb,0.52) (geom-fb,0.57)}; \addlegendentry{single-shot}
  \addplot[draw=cAccent!60!black,fill=cAccent!25] coordinates {(VLM-fb,0.62) (geom-fb,0.15)}; \addlegendentry{reviewed}
\end{axis}
\end{tikzpicture}
\caption{\textbf{An ungrounded reviewer makes layout worse (controlled ablation, two modalities).} Same tasks and proposer; only the reviewer's grounding varies. A VLM-feedback reviewer (\textsf{VLM-fb}) \emph{increases} mean geometric defects on both slides ($1.89\to2.44$) and figures ($0.52\to0.62$); the deterministic geometric-feedback reviewer (\textsf{geom-fb}) reduces them ($1.25\to0.33$; $0.57\to0.15$). The grounded signal is necessary on the feedback side, not just the metric side.}
\label{fig:ablation}
\end{figure}

\paragraph{Two modalities where the honest answer is ``no headroom.''} \emph{Video editing}, scored by Gemini~3.1~Pro's native full-clip rubric (a near-grounded instrument that observes pixels, not a still), is already strong single-shot ($0.70$, $22/45$ task-runs perfect) so the reviewed lift is small and not significant ($+0.04$); a hardened multi-step suite de-saturates it (single-shot ``fully correct'' $49\%\to11\%$). \emph{Deep research}, scored against exact provided facts, saturates ($0.982$, $+0.018$, n.s.): a frontier model knows well-documented facts and does not take the planted traps, so the Type-3d lift is bounded by the near-zero single-shot error rate---a genuine limit of the factual modality, since de-saturating it would require facts the judge itself cannot grade. We report these as scoped negatives rather than headline lifts.
