\section{Related Work}
\label{sec:related}

\paragraph{Test-time compute scaling.} Two axes are established. \emph{Reasoning} scaling (chain-of-thought, tree search, and extended thinking \citep{wei2022chain,yao2023tree,deepseek2025r1,snell2024scaling,kimi2025k15}) spends more tokens before committing. \emph{Sampling} scaling (self-consistency and best-of-$N$ \citep{wang2023selfconsistency,brown2024large}) draws multiple attempts and selects. Both transform the same frozen weights and prompt. We make precise (\Cref{sec:framework}) the sense in which they are bounded, and position \emph{interaction} as a third axis that imports external information.

\paragraph{Self-correction and critique.} Iterative refinement with model-generated feedback (Reflexion \citep{shinn2023reflexion}, ReAct \citep{yao2023react}, Self-Refine \citep{madaan2023selfrefine}, CRITIC \citep{gou2024critic}) is widely used, but its reliability is contested: \citet{huang2024selfcorrect} show models often cannot self-correct reasoning without an external signal, and \citet{kamoi2025correct} chart when correction does and does not help. Our taxonomy makes the distinction explicit. \emph{Ungrounded} critique (Type 1) is bounded, whereas refinement against an executed/measured result (Type 3) is not. And our in-modality control (\Cref{sec:interaction}) shows it is the external execution, not the extra critique pass, that carries the gain.

\paragraph{Grounded and execution feedback.} Tool use and execution feedback (Toolformer \citep{schick2023toolformer}, Gorilla \citep{patil2024gorilla}, Voyager \citep{wang2023voyager}, RLEF \citep{gehring2025rlef}, and, for website generation specifically, WebGen-Agent's multi-level visual/functional feedback \citep{webgen2025}), together with the thinking-vs-doing analysis of \citet{shen2025thinking}, establish that acting on an environment helps. We contribute the information-theoretic account of \emph{why}, a feedback-actionability taxonomy that predicts \emph{when}, and, crucially, the requirement that the \emph{evaluation} be grounded too. We also compare harness architectures (single-agent loops \citep{trankiela2026singleagent}, sample-and-select \citep{ruan2025iad}, proposer--reviewer) under a matched budget.

\paragraph{Model-as-judge reliability.} LLM- and VLM-as-judge evaluation is now standard \citep{zheng2023judging}, and its failure modes (position, verbosity, and self-preference biases, and broader reward-model fragility \citep{casper2023open}) are documented. Our finding is sharper and, to our knowledge, not previously isolated: for artifacts whose quality is a measurable \emph{physical} property (the geometry of a rendered layout), a screenshot judge is not merely biased but \emph{structurally blind} to defects cropped off-frame or below its acuity, so it both fails to score the defect and, used as a reviewer, fails to fix it. This is why a deterministic instrument is necessary, not merely preferable, to detect interaction scaling on visual modalities.

\paragraph{Practitioner evidence at scale.} Industrial deployment independently corroborates the gap our framework targets. \citet{hong2026aicoding} reports a controlled study at ByteDance ($3$ frontier coding models $\times$ $3$ agent frameworks $\times$ $100$ runs, $900$ total) on a single real product feature under a fixed prompt. Functional \emph{correctness} exceeded $80\%$ for every model--framework pair, yet ``deliverability'' axes (UI usability, reliability, maintainability, performance, compatibility) collapsed to a failing $40$--$60\%$ with high run-to-run randomness. Only after adding harness ``infrastructure'' (context engineering, architectural constraints, and persisted team memory) did deliverability rise to ${\sim}80\%$, while correctness barely moved ($80\%\!\to\!90\%$). The same report notes that despite ${>}90\%$ of one team's code being model-written, throughput rose only $1.6\times$, far short of the raw generation speedup. That shortfall is the macroscopic shadow of the quality single-shot generation does not deliver. This is the signature we formalize and measure: model and sampling scaling saturate the easy-to-see axis, while the interaction loop carries the rest. Tellingly, the report had to define a multi-axis quality metric before the gap was visible at all, an industrial analogue of our grounded-evaluation requirement. Its grounded fixes (e.g.\ browser-driven validation before commit) are Type-3 feedback in our taxonomy, not ungrounded self-review.

\paragraph{Information theory.} The data-processing inequality \citep{cover2006elements} is the lens for our bound: any post-processing of the model's outputs (reasoning, ungrounded review) cannot increase information about the target, whereas an environment observation introduces a new channel.
