\section{Discussion and Conclusion}
\label{sec:discussion}

\paragraph{Grounding is the load-bearing variable.} Two of the three test-time axes are internal and hit the internal ceiling because they never leave the model's token space. The third is not bounded by that ceiling, because a grounded instrument imports a real observation each cycle---it is of course still bounded, by what the instrument can observe and by the proposer's ability to act on it, which is exactly the coverage principle. We have argued and shown that grounding must hold on \emph{both} sides of the loop. On the feedback side, swapping only the reviewer's instrument shows improvement tracks coverage: execution feedback converges far cheaper than critique on behavioral bugs, a linter buys nothing on them, and a screenshot-fed VLM reviewer actively regresses slide layout. On the evaluation side, the default VLM-on-a-screenshot metric is structurally blind to the geometric defects the loop fixes, so the entire visual-modality effect is invisible until the metric is grounded in deterministic DOM geometry---and once it is, grounded feedback removes a large, decisive fraction of real layout defects on all four modalities (\Cref{fig:geomreduction}).

\paragraph{A methodological warning for the field.} Much of the recent excitement about ``LLMs that critique and improve their own visual output'' is measured with VLM judges. Our results imply many such measurements are uninformative on layout: the judge cannot see the defects, and using it as the reviewer can make things worse. Where an artifact's quality is a measurable physical property (geometry, execution, retrieval), the evaluation should be grounded in that property, not in a model's picture of it.

\paragraph{Grounding is the theory under ``loop engineering.''} Practitioners have converged on an informal art of \emph{loop engineering}, or \emph{loopcraft}---stacking an inner agent loop with an outer \emph{verification} loop that critiques and revises---as the design pattern for capable agents \citep{swyx2026loopcraft,langchain2026loopeng}. That verification loop is exactly the proposer--reviewer loop we study, and our results supply the variable the practitioner framing leaves implicit: \emph{grounding}, qualified by \emph{coverage}, decides when stacking such a loop pays off. A verification loop closed on a model's opinion---an LLM critique, or a VLM reading a screenshot---is subject to the internal ceiling (\Cref{sec:framework}) and, on visual artifacts, actively harmful (\Cref{sec:feedback}). Only a loop closed on an executed, rendered, or measured observation imports information and clears the ceiling. The lesson is not ``add more loops'' but ``ground the loop you add,'' on both the feedback and the evaluation side.

\paragraph{Limits, honestly stated.} Two modalities saturate: video editing is already strong single-shot (the lift is real only on a hardened multi-step suite), and deep research saturates because frontier models know well-documented facts and the judge cannot grade facts it does not have. Deterministic geometry applies to artifacts with a measurable intended layout (figures, slides, web, animations); for genuinely flowing content we still rely on a binary rubric, with the VLM-reliability caveat. Our static-form tier is confirmed only on its null prediction (no lift on bugs invisible in form), not yet on a positive one. And because the geometric instrument both supplies the feedback and scores the result, the reported reductions are partly an optimization of the metric itself; we argue in \Cref{sec:grounded_eval} (conservative thresholds, non-trivial magnitude, and the ungrounded-reviewer control that moves the same metric the wrong way) that they reflect real, human-visible defects, but a quantitative human-preference validation remains future work.

\paragraph{Future work and broader impact.} Two directions follow directly. First, deterministic instruments for modalities beyond layout and execution---differentiable or symbolic checks for audio, 3D, tabular, and UI-interaction artifacts---would extend grounded evaluation to where model judges are likewise blind. Second, the internal ceiling bounds an information channel; tightening it into a quantitative relationship between an instrument's coverage and achievable quality is open. On impact: grounding the metric is a cheap, deployable correction to a measurement error that, we argue, currently inflates reported progress on self-improving visual generation. The same instruments are auditable and reduce reliance on opaque model judges. The harness itself is scaffolding over a frozen model, so it inherits that model's capabilities and failure modes rather than introducing new ones.

\paragraph{Operational recipe.} Wrap a frozen frontier model in a proposer--reviewer harness whose instrument covers the defects that matter for the modality. For visual artifacts, ground \emph{both} the reviewer and the scorer in deterministic geometry: measure the rendered DOM, never a screenshot. Allocate the majority of the per-task budget to the proposer, and spend extra compute on samples rather than deeper iteration. For cost-sensitive deployment, distill the proposer into a small student---preserving its sampling variance---and run it inside the same harness. Interaction scaling is real, distinct from reasoning and sampling, and predictable from coverage. Once you ground the metric, it is plainly visible.
