\section{A Framework for Internal and External Test-Time Compute}
\label{sec:framework}

We do not propose a new self-correction method. We give an account of \emph{why} interaction is a separate scaling axis and \emph{when} a feedback channel helps (a coverage principle over a simple taxonomy), and we add a requirement prior work misses: the \emph{evaluation} must be grounded too, or the effect cannot be measured.

\subsection{Internal vs.\ external test-time compute}
\label{sec:internal-external}

Reasoning and sampling differ in mechanics, but not in kind. A longer chain of thought re-derives consequences of what the weights and prompt already contain; best-of-$N$ re-draws from the same distribution and picks one. In both, every token comes from the same two sources (frozen weights, fixed prompt), so we call them \textbf{internal} scaling. \textbf{External} scaling is different in kind: the artifact is handed to an instrument outside the model (a test runner, a layout engine, a search index), and the instrument's observation is fed back into the next generation. That observation reflects the \emph{artifact's actual behavior}, not the model's beliefs about it.

\subsection{A feedback taxonomy, and the coverage principle}
\label{sec:taxonomy}

We classify a reviewer's feedback channel by \emph{who produces the signal} (\Cref{fig:taxonomy}):
\begin{itemize}[leftmargin=1.4em,itemsep=1pt,topsep=2pt]
\item \textbf{Ungrounded feedback} is a model's opinion. An LLM re-reads its own artifact and critiques it; a VLM rates a screenshot. The screenshot case is worth care: a screenshot \emph{is} an instrument observation, but a lossy one that drops exactly the defects at issue (content off the canvas, overlaps too small to see), so the signal that actually enters the loop is the model's reading of those pixels. We classify a channel by what really produces its signal: here, the model.
\item \textbf{Grounded feedback} is an instrument's observation. Its subtypes are named by \emph{what the instrument observes}: \textbf{static form} (linters and type checkers inspect the source), \textbf{execution behavior} (tests run it and report the failing assertion), \textbf{rendered geometry} (a layout engine reports every element's true bounding box), \textbf{temporal behavior} (the same measurement per animation frame), and \textbf{external facts} (a search engine checks a claim against the world).
\end{itemize}
Grounding alone, however, is not sufficient, and this is the second half of the taxonomy:

\begin{center}
\begin{tikzpicture}
\node[draw=cGnd!70,fill=cGnd!5,rounded corners=3pt,inner sep=8pt,text width=0.9\linewidth]{%
\textbf{The coverage principle.} A feedback channel helps exactly as far as its instrument's observational reach. A linter is grounded but observes only form, so it cannot see a runtime bug; a screenshot is an observation whose channel crops the defect out, so a VLM judging it cannot see broken geometry. The instrument must \emph{observe the property that is broken}.};
\end{tikzpicture}
\end{center}

The coverage principle is what unifies the paper's results: it predicts that execution feedback beats critique on behavioral bugs while a linter does not (\Cref{sec:feedback}), that a screenshot-fed VLM fails both as a reviewer (\Cref{sec:feedback}) and as a judge (\Cref{sec:grounded_eval}), and that a deterministic geometry instrument succeeds at both.

\begin{figure}[t]
\centering
\resizebox{\linewidth}{!}{%
\begin{tikzpicture}[font=\small,
  ubox/.style={rounded corners=2pt,draw=cUng,fill=cUng!8,minimum height=10mm,align=center,inner sep=5pt},
  gbox/.style={rounded corners=2pt,draw=cGnd,fill=cGnd!8,minimum height=10mm,align=center,inner sep=5pt},
  lab/.style={font=\footnotesize\bfseries}]
  % ---- ungrounded lane (top) ----
  \node[ubox] (crit) at (0,0) {LLM critique\\\footnotesize re-reads its own output};
  \node[ubox, right=5mm of crit] (vlmshot) {VLM on a screenshot\\\footnotesize channel drops the defect};
  \node[font=\footnotesize\itshape,cUng!85!black,right=8mm of vlmshot,align=left]
    {bounded by what the model already knows\\(the internal ceiling, \Cref{sec:ceiling})};
  % ---- grounded lane (bottom) ----
  \node[gbox] (lint) at (0,-3.1) {static form\\\footnotesize linter, type checker};
  \node[gbox, right=4mm of lint] (exec) {execution\\\footnotesize tests, tracebacks};
  \node[gbox, right=4mm of exec] (geom) {rendered geometry\\\footnotesize DOM bounding boxes};
  \node[gbox, right=4mm of geom] (temp) {temporal\\\footnotesize per-frame geometry};
  \node[gbox, right=4mm of temp] (fact) {external facts\\\footnotesize search verification};
  % coverage arrow under grounded row
  \draw[-{Latex[length=2.2mm]},thick,cGnd!70!black]
    ([yshift=-3.5mm]lint.south west) -- ([yshift=-3.5mm]fact.south east)
    node[midway,below=2pt,font=\footnotesize,cGnd!70!black]
    {coverage, what the instrument can observe: \ form $\rightarrow$ behavior $\rightarrow$ world};
  \begin{scope}[on background layer]
    \node[draw=cUng!60,dashed,rounded corners,fill=cUng!3,fit=(crit)(vlmshot),inner sep=2.5mm,
      label={[cUng!85!black,lab]above:UNGROUNDED: a model's opinion}] (ug) {};
    \node[draw=cGnd!60,dashed,rounded corners,fill=cGnd!3,fit=(lint)(fact),inner sep=2.5mm,yshift=-2mm,
      label={[cGnd,lab]above:{GROUNDED: an instrument's observation \normalfont\itshape imports new information, as far as it can observe}}] (gd) {};
  \end{scope}
\end{tikzpicture}}
\caption{\textbf{The feedback taxonomy.} Feedback is either a model's opinion (\emph{ungrounded}, top) or an instrument's observation of the artifact (\emph{grounded}, bottom); grounded subtypes are named by what the instrument observes, and the \emph{coverage} axis orders them by reach. The screenshot-fed VLM sits in the ungrounded lane because the signal that enters the loop is a model's reading of an already-lossy view. Ungrounded feedback is subject to the internal ceiling (\Cref{sec:ceiling}); grounded feedback escapes it, but only for defects inside the instrument's coverage: a linter cannot see a runtime bug, and a screenshot cannot see off-canvas overflow.}
\label{fig:taxonomy}
\end{figure}

\subsection{Why internal scaling saturates}
\label{sec:ceiling}

\begin{figure}[t]
\centering
\begin{tikzpicture}
\begin{axis}[width=0.86\linewidth,height=5.6cm,
  xmode=log,log basis x=10,
  xlabel={per-task token budget $B$},ylabel={code pass rate (\%)},
  xmin=820,xmax=26000,ymin=54,ymax=104,clip=false,
  xtick={1000,5000,20000},xticklabels={1K,5K,20K},
  ytick={60,70,80,90,100},ymajorgrids,
  legend style={at={(0.5,-0.30)},anchor=north,legend columns=2,
    /tikz/every even column/.append style={column sep=0.5cm}},legend cell align=left]
  % region only interaction reaches (above the oracle-sampling ceiling)
  \fill[cGnd!8] (axis cs:820,86.7) rectangle (axis cs:26000,104);
  \node[font=\footnotesize\itshape,cGnd!80!black,anchor=east] at (axis cs:13500,102.3)
     {reached only by interaction};
  \draw[cMut,densely dotted] (axis cs:820,86.7)--(axis cs:26000,86.7);
  \node[font=\footnotesize,cMut,anchor=south west] at (axis cs:870,86.9)
     {internal ceiling (oracle best-of-$N$)};
  \addplot[cGnd,line width=1.4pt,mark=*,mark size=1.9,mark options={fill=cGnd,draw=white,line width=0.5pt}]
    coordinates {(1000,62.2)(5000,91.1)(20000,100.0)};
    \addlegendentry{proposer--reviewer \emph{(external)}}
  \addplot[cGndDk,line width=1.1pt,mark=square*,mark size=1.7,mark options={fill=cGndDk,draw=white,line width=0.5pt}]
    coordinates {(1000,57.8)(5000,93.3)(20000,97.8)};
    \addlegendentry{single-agent loop \emph{(external)}}
  \addplot[cUng,line width=1.1pt,mark=triangle*,mark size=2.2,mark options={fill=cUng,draw=white,line width=0.5pt}]
    coordinates {(1000,80.0)(5000,77.8)(20000,86.7)};
    \addlegendentry{best-of-$N$, oracle verifier \emph{(internal)}}
  \addplot[cUngDk,line width=1.1pt,mark=diamond*,mark size=2.2,mark options={fill=cUngDk,draw=white,line width=0.5pt}]
    coordinates {(1000,60.0)(5000,73.3)(20000,73.3)};
    \addlegendentry{reasoning-only \emph{(internal)}}
  % selective end labels: the two ceilings and the harness endpoint
  \node[font=\scriptsize\bfseries,cGnd,anchor=south east,inner sep=1pt] at (axis cs:20000,100.4) {100.0};
  \node[font=\scriptsize,cUngDk,anchor=north west,inner sep=2pt] at (axis cs:20000,73.0) {73.3};
\end{axis}
\end{tikzpicture}
\caption{\textbf{Internal scaling saturates; external scaling does not (P1).} Code pass rate vs.\ per-task token budget on the $15$ hard tasks ($3$-seed means; full table in \Cref{tab:scaling-curves}). Both internal strategies flatten as the budget grows: reasoning-only at its information ceiling, and best-of-$N$ at $86.7\%$ \emph{even with an oracle (ground-truth-test) verifier}, capped by the high-probability part of the model's own output distribution. The two external strategies import execution feedback each cycle and climb past that ceiling, the proposer--reviewer harness to a strict $100\%$ with zero seed variance. We run this experiment in \Cref{sec:interaction}.}
\label{fig:scaling}
\end{figure}

The reason is one sentence long: \emph{a model re-reading its own output cannot learn anything it did not already know.} A critique produced by the same weights that produced the artifact is post-processing, and post-processing cannot add information about the correct answer. Sampling is bounded for an equally simple reason: best-of-$N$ can only \emph{select} among the candidates the model actually draws, so even a perfect verifier cannot return a program the model was never going to write. Longer thinking rearranges the known; more samples explore the already-likely. We call the resulting plateau the \textbf{internal ceiling}. The formal version (three lines of data-processing inequality) is deferred to \Cref{app:ceiling}; since the inequality bounds an information channel rather than achievable quality directly, the paper's evidence is experimental. \Cref{fig:scaling} makes this concrete: on hard code, reasoning-only and \emph{oracle-verified} best-of-$N$ flatten as the budget grows, while interaction strategies climb past them to a perfect pass rate.

Grounded feedback escapes the ceiling because the observation is computed from the artifact's real behavior (a failing assertion, a measured bounding box, a search verdict), and conditioning the \emph{next} proposal on it lets the model synthesize a candidate it would never have sampled unaided. This is the difference from selection (which cannot create a missing candidate) and from reasoning (which adds no information): interaction is a new information channel, not a constant-factor speedup.

\paragraph{The symmetric claim: an ungrounded metric cannot measure the gain.} The same argument applies to the \emph{scorer}. A judge that is itself a model looking at a lossy view of the artifact (a VLM reading a cropped screenshot) cannot certify an improvement that lives in the part of the artifact its channel drops. Grounding the evaluation is therefore not polish but a precondition for \emph{detecting} interaction scaling wherever quality is not fully visible to the default judge. \Cref{sec:grounded_eval} shows this is exactly the situation for rendered visual artifacts.

\subsection{Three testable predictions}
\label{sec:predictions}

\textbf{(P1)} At a matched token budget, reasoning-only and best-of-$N$ scaling saturate strictly below what interaction reaches (\Cref{sec:interaction}). \textbf{(P2)} Holding the task suite fixed and swapping only the reviewer's feedback channel, improvement tracks the instrument's \emph{coverage} of the defects present, not the act of reviewing (\Cref{sec:feedback}). \textbf{(P3)} On modalities whose quality is invisible to the default judge, the gain is only detectable with a grounded metric, and an ungrounded reviewer can even regress quality (\Cref{sec:feedback,sec:grounded_eval}).
