\section{Internalizing the Harness into a Small Student}
\label{sec:distill}

Two things could carry the harness's gains: the grounded scaffolding, or behavior absorbed into the weights. This section asks how much is the latter: whether the loop's interaction-quality behavior can be \emph{internalized} so a small model emits high-quality artifacts in fewer passes. It is a complement to the thesis, not a substitute: the cheaper internalized proposer is still meant to run \emph{inside} the same grounded harness, so internalization and grounding compose. We distill judge-filtered teacher trajectories from the harness into an $8$B Qwen3-VL student via SFT.

\paragraph{Result.} The student recovers about half the teacher's mean@1 capability on a $44$-task out-of-distribution suite, rising with a second sample (\Cref{tab:distill-ood}), and passes a substantial fraction of an $18$-task hard held-out suite (\Cref{tab:distill-headline}), at roughly $10\times$ lower deployment cost than running the harness over a frontier model (\Cref{fig:distill}). Run back inside the harness, the student recovers still more of the gap.

\begin{figure}[t]
\centering
\begin{tikzpicture}
\begin{axis}[width=0.66\linewidth,height=4.4cm,clip=false,
  ybar,bar width=30pt,ymin=0,ymax=1.2,
  ylabel={capability vs.\ teacher ($\times$)},
  symbolic x coords={mean@1,pass@2},xtick=data,enlarge x limits=0.6,
  nodes near coords={\pgfmathprintnumber[fixed,fixed zerofill,precision=2]{\pgfplotspointmeta}$\times$},
  nodes near coords style={font=\footnotesize\bfseries,anchor=south},
  ymajorgrids,ytick={0,0.25,0.5,0.75,1.0},yticklabel={\pgfmathprintnumber{\tick}$\times$}]
  % teacher reference band spanning the full plot width
  \fill[cGrey!7] ({rel axis cs:0,0}|-{axis cs:mean@1,1.0}) rectangle ({rel axis cs:1,0}|-{axis cs:mean@1,1.2});
  \draw[cGrey,densely dashed,thick] ({rel axis cs:0,0}|-{axis cs:mean@1,1.0}) -- ({rel axis cs:1,0}|-{axis cs:mean@1,1.0});
  \node[font=\footnotesize,cGrey,anchor=south east] at ({rel axis cs:0.99,0}|-{axis cs:mean@1,1.01}) {teacher ($1\times$)};
  \addplot[draw=cAccent!60!black,fill=cAccent!30] coordinates {(mean@1,0.51) (pass@2,0.70)};
  % sampling-lift annotation in the empty band between the bar tops and the teacher line
  \node[font=\scriptsize,cRV!45!black,anchor=center]
    at ($(axis cs:mean@1,0.90)!0.5!(axis cs:pass@2,0.90)$)
    {sampling adds $+0.19\times$};
  \legend{}
\end{axis}
\end{tikzpicture}
\caption{\textbf{An $8$B student internalizes most of the teacher's interaction quality.} Distilled $8$B markdown student as a fraction of the frontier teacher on the $44$-task OOD suite (\Cref{tab:distill-ood}): a single sample recovers $0.51\times$ and two samples $0.70\times$ of teacher capability (sampling closes the gap), at ${\sim}10\times$ lower deployment cost. On the $18$-task hard held-out suite the same student passes $44\%$/$56\%$ at pass@$1$/pass@$3$.}
\label{fig:distill}
\end{figure}

\paragraph{Variance is a budget (a cautionary finding).} Reinforcement fine-tuning (RFT) on top of SFT improves every per-turn consistency metric we measure \emph{and} regresses pass@3 (\Cref{tab:distill-rft}). For sampling-based deployment, output-distribution variance \emph{is} the inference-scaling lever, and any post-training that reduces it spends from the same budget best-of-$N$ draws on. Variance is not noise to minimize. It is the resource sampling converts into quality. The operational recipe is therefore: SFT to internalize interaction quality, keep the proposer's sampling temperature, and run it inside the grounded harness.
