\section{Robustness}
\label{sec:robustness}

\paragraph{Cross-model, code.} The effect is architectural, not a Claude artifact. On the $15$ hard code tasks, over \emph{three} independent on-policy seeds per family, the harness lifts Sonnet~4 by $+33.3$\,pp, Qwen3-235B by $+22.2$\,pp, and GPT-5 by $+20.0$\,pp (on a higher single-shot baseline; \Cref{fig:crossmodel}). The reviewed ceiling has \emph{zero} seed variance for every family (Sonnet $100.0$, GPT-5 $100.0$, Qwen $93.3\%$, all $\pm0.0$\,pp) while single-shot fluctuates seed-to-seed: the grounded loop converges to the same ceiling regardless of the starting draw. All three families benefit from the same execution-grounded loop.

\begin{figure}[t]
\centering
\begin{tikzpicture}
\begin{axis}[width=0.82\linewidth,height=3.8cm,
  ybar,bar width=11pt,ymin=50,ymax=104,
  ylabel={code pass-rate (\%)},
  symbolic x coords={Sonnet 4,Qwen3-235B,GPT-5},xtick=data,
  enlarge x limits=0.28,
  nodes near coords,nodes near coords style={font=\footnotesize,/pgf/number format/.cd,fixed,precision=1},
  ymajorgrids,ytick={60,80,100},
  legend style={font=\footnotesize,at={(0.5,1.02)},anchor=south,legend columns=2,draw=cGrey!50}]
  \addplot[draw=cSS!60!black,fill=cSS!30,error bars/.cd,y dir=both,y explicit]
    coordinates {(Sonnet 4,66.7) +- (0,6.7) (Qwen3-235B,71.1) +- (0,3.8) (GPT-5,80.0) +- (0,6.7)}; \addlegendentry{single-shot}
  \addplot[draw=cRV!60!black,fill=cRV!35,error bars/.cd,y dir=both,y explicit]
    coordinates {(Sonnet 4,100) +- (0,0) (Qwen3-235B,93.3) +- (0,0) (GPT-5,100) +- (0,0)}; \addlegendentry{reviewed}
\end{axis}
\end{tikzpicture}
\caption{\textbf{The execution-grounded loop helps every model family.} Single-shot vs.\ harness-reviewed code pass-rate for three model families, each over three independent on-policy seeds ($+33.3$/$+22.2$/$+20.0$\,pp). Whiskers are $\pm1$ SD across seeds; the reviewed bars carry \emph{zero} seed variance for all three families.}
\label{fig:crossmodel}
\end{figure}

\paragraph{Cross-model, grounded geometry.} To rule out a Claude-specific or prompt-specific explanation of the grounded-evaluation effect, we run the same deterministic geometric-feedback harness with \textbf{Gemini~3.1~Pro} as the proposer on the dense real-paper slide suite, scored by the identical model-free DOM instrument. The effect reproduces and is in fact larger: mean defects fall $3.50\to0.25$ ($-93\%$) over three seeds ($n{=}36$), geometrically-clean slides rise $17/36\to34/36$, and $19$ of $20$ decisive task-runs improve (sign-test $p=4\times10^{-5}$). Gemini's single-shot slides are \emph{more} defect-prone than Sonnet's ($3.50$ vs.\ $1.25$), and the grounded loop removes almost all of it. Because the instrument depends on no model, this rules out a Claude-specific or scorer-specific explanation.

\paragraph{Held-out generalization.} On a $32$-task held-out code suite constructed after the method was fixed, the harness lifts $90.6\%\to100\%$, recovering $3/3$ first-shot failures with zero regressions---no tuning to the original tasks.

\paragraph{Budget allocation.} Sweeping the proposer/reviewer token split on a fixed budget produces an $86.6$\,pp spread in pass-rate, monotone in the proposer's per-call share: the optimum is \emph{propose-heavy} (allocate the majority to generation, reserve enough for the reviewer to localize defects). Extra inference compute is better spent on additional samples than on deeper iteration past the early-stop point.
