\section{Setup: A Budget-Aware Proposer--Reviewer Harness}
\label{sec:method}

\paragraph{Architecture.} The harness is pure scaffolding around a \emph{frozen} frontier model (Claude Sonnet~4, \texttt{claude-sonnet-4-20250514}, temperature~$0$ unless noted): no fine-tuning, no retrieval, no learned controller. A \textbf{proposer} produces an artifact. An \textbf{instrument} executes, renders, or measures it to produce a grounded signal. A \textbf{reviewer} (the same model in a diagnostic role) turns that signal into a structured list of defects, and the proposer revises. The loop repeats up to an iteration cap and \emph{keeps the best-scoring iteration}, so the reviewed result can never score below single-shot under the same metric. Single-shot is just one proposer call.

\paragraph{The three grounded instruments.} Every modality is paired with a deterministic or near-deterministic instrument that both \emph{feeds back} and \emph{scores}:
\begin{itemize}[leftmargin=1.4em,itemsep=1pt,topsep=2pt]
\item \textbf{Execution} (\texttt{pytest}): exact pass/fail plus the failing assertion and traceback. Each task ships a multi-assertion suite targeting the edge cases that separate correct from plausible-but-wrong (e.g.\ the semantic-version comparator is checked on pre-release precedence), and the deep-spec suite is additionally validated against a reference implementation. A ``pass'' therefore certifies the specified behavior, and the oracle best-of-$N$ baseline of \Cref{sec:interaction} selects against these \emph{same} tests, so its ceiling is measured by the identical criterion.
\item \textbf{Rendered geometry}: the artifact is rendered headless and every element's true bounding box is read from the layout engine. We compute, exactly: text-on-text overlap ($\ge 6$\,px), out-of-bounds/clipping, container overflow, document overflow ($>16$\,px), and \emph{box-group misalignment} (rows or columns of card-like boxes flagged for unequal size, misaligned far edges, or uneven gutters). Web pages are scored at desktop and mobile widths; animations are probed at each sampled frame.
\item \textbf{Native video} (Gemini~3.1~Pro): the rendered clip is judged whole against per-requirement binary checks, a near-grounded substitute where pixels, not a still, are observed.
\end{itemize}
Where no deterministic instrument exists (flowing web content, factual research), we fall back to a binary per-requirement rubric, flagged as such.

\begin{table}[t]
\centering
\small
\caption{Modality instantiations. Each modality pairs a grounded feedback channel (named per the taxonomy of \Cref{fig:taxonomy}) with the instrument that both drives revision and scores the result. ``$\dagger$'' marks the four modalities scored by the deterministic DOM-geometry instrument and analyzed jointly in \Cref{sec:grounded_eval}.}
\label{tab:setup}
\begin{tabular}{llll}
\toprule
Modality & Feedback channel & Grounded instrument & Hard suite ($N$) \\
\midrule
Code             & execution     & \texttt{pytest} pass/fail + traceback   & $15$ (+$11$ deep-spec) \\
Academic figures$^\dagger$ & geometry & DOM geometry + alignment           & $20$ \\
Slides$^\dagger$           & geometry & DOM geometry + alignment           & $12$ real-paper \\
Web pages$^\dagger$        & geometry & DOM geometry, multi-width          & $20$ \\
Animations$^\dagger$       & temporal & DOM geometry, per-frame            & $20$ SVG/CSS \\
Video editing    & exec.\ + temporal & script exec + Gemini-native rubric & $15$ \\
Deep research    & external facts & search-grounded factual rubric         & $15$ \\
\bottomrule
\end{tabular}
\end{table}

All artifacts are produced under a common \emph{design-principle} generation prompt (proximity, alignment, repetition, contrast, deliberate color) so that single-shot quality is already strong: the headroom we measure reflects genuine task difficulty, not a weak prompt.

\paragraph{Budget and protocol.} A run has a token budget $B$ split across proposer and reviewer calls; \Cref{sec:robustness} sweeps the split, and \Cref{tab:setup} lists the per-modality instruments. Unless noted, the proposer runs at temperature $0$, and each modality is evaluated over three independent seeds, keeping the best-scoring iteration (cap $\le 3$ for geometry, $\le 5$ for code); we compare single-shot against reviewed on each (task, seed) pair. For each result we report a two-sided paired sign test over the decisive pairs and, for the geometry and code effects, a paired-bootstrap $95\%$ confidence interval on the effect size; exact values accompany each figure.
