\section{The Active Ingredient Is Execution, Not Review}
\label{sec:channel}

\paragraph{Code as a clean case study.} Code is the cleanest demonstration because the channel is binary and the score is objective. Single-shot pass-rate is $66.7\%$ ($10/15$); the harness lifts it to a strict $100\%$ in all three runs, recovering every first-shot failure with zero regressions. The mechanism is consistent across recovered traces: turn~1 emits code that compiles but mishandles an edge case (off-by-one boundaries, semantic-version comparison, CIDR arithmetic); turn~2 receives the \texttt{pytest} traceback---the failing assertion, the offending input, expected vs.\ actual---and rewrites the localized block; turn~3 confirms. Reasoning alone cannot supply this: the model has no way to know its boundary condition is wrong until a concrete test exercises it. A second, harder $11$-task suite of deep-spec implementations (a from-scratch JSON parser, an A1 spreadsheet-reference resolver, a minimal edit-script diff, an arithmetic/boolean expression evaluator, $\dots$; each validated against a reference solution that passes its own tests) corroborates this at larger headroom: single-shot $69.7\%\to$ strict $100\%$, $+30.3$\,pp (bootstrap 95\% CI $[15.2,45.5]$\,pp), all $10$ failing task-runs recovered, $0$ regressions (sign-test $p=0.002$).

\paragraph{An in-modality control isolates execution (P2).} The cross-modal ordering could be confounded by task difficulty, so we fix the task suite (the $15$ hard code tasks) and vary \emph{only} the feedback channel: a Type~1 reviewer (LLM cross-review, no execution), a Type~2 reviewer (LLM + \texttt{ruff} static analysis, still no execution), and the Type~3a baseline. Forbidden from seeing test results, the Type~1/2 reviewers reach a reviewed ceiling of $13/15$; the Type~3a (execution) reviewer reaches $14/15$ and does so $\sim$$2.5\times$ more token-efficiently ($1.53$ vs.\ $2.07$ iterations). The separation is not the extra reviewing pass---both have one---but the \emph{execution} of the tests. (This Type~3a \emph{control} is a single leaner-budget reference seed and caps at $14/15$; its lone residual is \texttt{code\_011}, CJK text wrapping, which the full proposer--reviewer harness at the $20$K budget recovers in all three seeds---hence the strict $100\%$ headline above.) (Type~1 and Type~2 do not separate because the seeded bugs are runtime-logic errors that \texttt{ruff} cannot see; a \texttt{mypy}/\texttt{semgrep}-catchable bug set is where the strict Type~2~$>$~Type~1 prediction should be tested.)
